---
title: "Homework 2"
subtitle: "Research in Health Economics, Spring 2025"
author: "Megan Zheng"
format:
    pdf: 
        output-file: "zheng-m-hwk2-3"
        output-ext: "pdf"
        header-includes: 
            - \usepackage{float}
            - \floatplacement{table}{H}
knitr:
    opts_chunk:
        warning: false

---

```{r}
#| include: false
#| eval: true
library(tidyverse)
load("hwk2_workspace.Rdata")
```

```{r}
#| include: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra)
               
```

In this assignment, I wil be analyzing the HCRIS data to pull informtion from their reports and estimate ATEs. The analysis and the raw data can be found in another script/document.

The GitHub repository for this work is available [here](https://github.com/meganesmeralda/homework2). Enjoy!

\newpage
# Summarize the Data

\noindent Question 1. How many hospitals filed more than one report in the same year? Show your answer as a line graph of the number of hospitals over time.<br>

```{r}
#| echo: false
report_counts <- duplicate.hcris %>%
  group_by(fyear) %>%
  summarise(num_hospitals = n_distinct(provider_number))

q1 <- ggplot(report_counts, aes(x = fyear, y = num_hospitals)) +
  geom_line() +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),  
    axis.title = element_text(size = 12),  
    axis.text = element_text(size = 10),  
    panel.grid = element_blank()) +
   labs(
    title = "Number of Hospitals Filing More than One Report per Year",
    x = "Fiscal Year",
    y = "Number of Hospitals")

q1
``` 
\vspace{1cm}

\newpage
\noindent Question 2. After removing/combining multiple reports, how many unique hospital IDs exist in the data?

```{r}
#| echo: false
unique_hospital_count <- final.hcris.data %>%
  group_by(year) %>%
  summarise(num_unique_providers = n_distinct(provider_number), .groups = 'drop')

library(knitr)

kable(unique_hospital_count, 
      col.names = c("Year", "Unique Providers"),
      caption = "Unique Hospital IDs per Year")
```

\newpage

\noindent Question 3. What is the distribution of total charges in each year? Show your results with a “violin” plot.

```{r}
#| echo: false
q3 <- ggplot(final.hcris.data, aes(x = fyear, y = log(tot_charges))) +
  geom_violin(fill = "pink", color = "darkblue") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),  
    axis.title = element_text(size = 10),  
    axis.text = element_text(size = 5),  
    panel.grid = element_blank()) +
   labs(
    title = "Distribution of Total Charges by Year",
    x = "Year",
    y = "Log of Total Charges")

q3
```

\newpage
\noindent Question 4. What is the distribution of estimated prices in each year? Present your results with a violin plot.

```{r}
#| echo: false
final.hcris.data <- final.hcris.data %>%
    mutate(discount_factor = 1 - tot_discounts / tot_charges)

final.hcris.data <- final.hcris.data %>%
    mutate(price_num = (ip_charges + icu_charges + ancillary_charges) * discount_factor - tot_mcare_payment)

final.hcris.data <- final.hcris.data %>%
    mutate(price_denom = tot_discharges - mcare_discharges)

final.hcris.data <- final.hcris.data %>%
    mutate(price = price_num / price_denom)


# Perform the filtering operation
final.hcris.data <- final.hcris.data %>%
    filter(price >= quantiles[1], price <= quantiles[2])

# Print the number of rows after filtering
print(nrow(final.hcris.data))

quantiles <- quantile(final.hcris.data$price, c(0.01, 0.99), na.rm = TRUE)
final.hcris.data <- final.hcris.data %>%
    filter(price >= quantiles[1], price <= quantiles[2])

final.hcris.data$fyear <- as.factor(final.hcris.data$fyear)

q4 = ggplot(final.hcris.data, aes(x = fyear, y = price)) +
    geom_violin(fill = "lightblue", color = "darkblue") +  
    labs(
        title = "Distribution of Estimated Prices by Year",
        x = "Year",
        y = "Estimated Price"
    ) +
    theme_minimal() +  
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

q4
```
\newpage

# Estimating ATEs

\noindent Question 5. Calculate the average price among penalized versus non-penalized hospitals.

From my analysis, the average price among penalized hospitals is $9685 and $9323 among non-penalized hospitals. 

\newpage
\noindent Question 6. Split hospitals into quartiles based on bed size. Provide a table of the average price among treated/control groups for each quartile.

```{r}
#| echo: false
library(knitr)

kable(quartile_summary, 
      col.names = c("Bed Size", "No Penalty", "Penalty"),
      caption = "Average Prices by Bed Quartile and Penalty")
```

\newpage
\noindent Question 7. Find the average treatment effect using each of the following estimators, and present your results in a single table. 

```{r}
#| echo: false
library(knitr)

kable(q7, 
      col.names = c("Method", "ATE", "SE"),
      caption = "Average Treatment Effect Estimates")
```

\newpage
# Question 8
With these different treatment effect estimators, are the results similar, identical, very different?


\newpage
# Question 9
Do you think you’ve estimated a causal effect of the penalty? Why or why not? (just a couple of sentences)


\newpage
# Question 10
Briefly describe your experience working with these data (just a few sentences). Tell me one thing you learned and one thing that really aggravated or surprised you.